---
title: "China mainland population density"
date: "2016-07-14T13:39:46+02:00"
author: "Datapleth.io"
tags: ["china", "map", "demography"]
categories: ["dataviz","geography"]
banner: "blog/2016_07_14_gridded_population_of_china_files/figure-html/mapDensity-1.png"
output: html_document
---

In an earlier
[post](https://datapleth.io/blog/2015/10/19/2015-10-19_china-urbanization/) we
mapped the urbanization rate of China at province level. In this post we will go
futher by visualizing where Chinese people are living using a gridded population
map.

We will use the NASA dataset (Population Density Grid, v3 (1990, 1995, 2000)) 
which consists of estimates of human population by 2.5 arc-minute grid cells. 

> A proportional allocation gridding algorithm, utilizing more than 300,000
national and sub-national administrative units, is used to assign population 
values to grid cells. The population density grids are derived by dividing the 
population count grids by the land area grid 
and **represent persons per square kilometer**.


## Required libraries

We need several R packages for this study.

```{r loadLibraries, message=FALSE, warning=FALSE}
library(maptools)
library(gtools)
library(rgdal)
library(raster)
library(rasterVis)
library(ggplot2)
library(ggthemes)
```


## Get and clean population and grid data


The data is available for free to download on SEDAC website [https://sedac.ciesin.columbia.edu/data/set/gpw-v3-population-density](https://sedac.ciesin.columbia.edu/data/set/gpw-v3-population-density)

Please note that the Trustees of Columbia University in the City of New York and
the Centro Internacional de Agricultura Tropical (CIAT) hold the copyright of 
this dataset. Users are prohibited from any commercial, non-free resale, or 
redistribution without explicit written permission from CIESIN or CIAT. Users 
should acknowledge CIESIN and CIAT as the source used in the creation of any
reports, publications, new data sets, derived products, or services resulting
from the use of this data 
set.

There are 3 formats available: .bil, .ascii or .grid
(https://en.wikipedia.org/wiki/Esri_grid)
We will use directly the ASCII version which can be read with `readAsciiGrid()`
from the `maptool` package.

We select data for China only and download 1 datasets : Population density grid;
We use a cached version stored on datapleth.io cloud and load the file raster.

```{r configureFileSource, message=FALSE, warning=FALSE}
# chnds00g	population densities in 2000, unadjusted, persons per square km
densGriCountFile <- "https://data.datapleth.io/ext/gridded-Population-of-China/chndens-ascii/chnds00g.asc"
download.file(
  url = densGriCountFile, 
  destfile = "./data/chnds00g.asc", 
  method = "wget", quiet = TRUE,
  mode = "w", cacheOK = TRUE
)
dens <- raster::raster("./data/chnds00g.asc")
fromDisk(dens)
```


## Visualise gridded data

```{r visualizeData}
plot(dens, main="China population density grid")
# Around Shanghai
plot(
  crop(
    dens,
    c(116, 123, 29, 33)
  ),
  main="Shanghai population density grid"
)
```


## Toward a map of China density 

We first load China map with regions 
(see [earlier post](https://datapleth.io/blog/2015/10/19/2015-10-19_china-urbanization/))

```{r baseMapChina, message=FALSE, warning=FALSE}
ChinaPolygonsLevel1 <- rgdal::readOGR("./data/CHN_adm1.shp")

## Fix English names, simplify
ChinaPolygonsLevel1@data$NAME_1 <- as.character(ChinaPolygonsLevel1@data$NAME_1)
ChinaPolygonsLevel1@data[grep("Xinjiang Uygur", ChinaPolygonsLevel1@data$NAME_1),]$NAME_1 <- "Xinjiang"
ChinaPolygonsLevel1@data[grep("Nei Mongol", ChinaPolygonsLevel1@data$NAME_1),]$NAME_1 <- "Nei Menggu"
ChinaPolygonsLevel1@data[grep("Ningxia Hui", ChinaPolygonsLevel1@data$NAME_1),]$NAME_1 <- "Ningxia"
ChinaPolygonsLevel1@data$NAME_1 <- as.factor(ChinaPolygonsLevel1@data$NAME_1)

# Use level1 as index & Province name as id
ChinaLevel1Data <- ChinaPolygonsLevel1@data
ChinaLevel1Data$id <- ChinaLevel1Data$NAME_1

# Fortify the data (polygon map as dataframe) using english names
ChinaLevel1dF <- fortify(ChinaPolygonsLevel1, region = "NAME_1")

## Merge polygons and associated data in one data frame by id (name of the province in chinese)
ChinaLevel1 <- merge(ChinaLevel1dF, ChinaLevel1Data, by = "id")
rm(ChinaPolygonsLevel1, ChinaLevel1dF, ChinaLevel1Data)
```

Then we convert the density raster data to data.frame in order to use
ggplot in the next steps.


```{r}
## convert the raster to points (to plot with ggplot)
raster_points_dens <- rasterToPoints(dens)
raster_points_dens <- data.frame(raster_points_dens)
colnames(raster_points_dens) <-c('x','y','density')
```

We make a first plot.

```{r message=FALSE, warning=FALSE}
md <- ggplot(data=raster_points_dens, aes(y=y, x=x)) 
md <- md + geom_raster(aes(fill=density))
md + scale_fill_gradient(low="white", high="blue")
```

The problem of this visualisation is the large extend of distribution of
densities. China is definitely a wide country which is dense and empty at the
same time ! So we switch to a log scale to get a better vizualisation.

```{r}
qplot(
  x = raster_points_dens$density,
  fill=..count.., 
  geom="histogram",
  bins = 200
)
qplot(
  x = log(raster_points_dens$density+1),
  fill=..count.., 
  geom="histogram",
  bins = 200
) 

```




## Map of China density grid

As described above, we use a log(x+1) scale for the color scale 
(using `trans="log1p"`) in the geom_raster.

```{r mapDensity}
g <- ggplot() + theme_tufte() +
  ## Projected shadows
  geom_polygon(
    data = ChinaLevel1,
    aes(x = long + 0.7, y = lat - 0.5, group = group),
    color = "grey50", size = 0.1, fill = "grey50"
  ) +
  ## Add Raster
  geom_raster(
    data=raster_points_dens,
    aes(y=y, x=x, fill=density)
  ) +
  scale_fill_distiller(
    palette = "Spectral",
    trans="log1p",
    breaks = c(0, 3, 10, 50, 200, 1000, 5000, 20000)
  ) +
  ## Province boundaries
  geom_polygon(
    data = ChinaLevel1,
    aes( x = long, y = lat, group = group, fill = NULL),
    alpha = 0.01, color = "black", size = 0.2
  ) +
  ## Styling
  labs(title = "China mainland - population density & provinces") +
  theme( axis.text=element_blank(), axis.ticks=element_blank(),
         axis.title=element_blank())
print(g)
```





## References

### Source 

Center for International Earth Science Information Network - CIESIN - Columbia 
University, United Nations Food and Agriculture Programme - FAO, and Centro 
Internacional de Agricultura Tropical - CIAT. 2005. Gridded Population of the 
World, Version 3 (GPWv3): Population Count Grid. Palisades, NY: NASA 
Socioeconomic Data and Applications Center (SEDAC). 
https://doi.org/10.7927/H4639MPP. Accessed `r Sys.Date()`.


### Similar work

http://metrocosm.com/world-population-split-in-half-map/

### Other example of use of `raster` package

https://pakillo.github.io/R-GIS-tutorial/#raster
http://zevross.com/blog/2015/03/30/map-and-analyze-raster-data-in-r/
https://jeffreybreen.wordpress.com/tag/raster/