---
title: "Getting and cleaning data, example of Chinese airports - Part 4/5"
date: "2016-01-01T13:39:46+02:00"
author: "Datapleth.io"
tags: ["airports", "data"]
categories: ["geography","china"]
banner: "blog/2016-01-01_China_airports_part_4_files/figure-html/mapAO2-2.png"
output: html_document
---


## Introduction
### Objective  
One of the big problem for anybody interested in China and data science is the 
availability of data sets. There are limited free resources available and they 
are often incomplete or inaccurate. Getting data and especially cleaning data 
becomes one of the biggest pain of data science applied to China.

The objective of this group of post is to illustrate the problem and associated
process on a specific example: plot a map of the airports of mainland China.

In this forth part, we merge the two data obtained in
[part 2](https://datapleth.io/blog/2015/12/30/2015-12-30_china_airports_part_2/)
and
[part 3](https://datapleth.io/blog/2015/12/31/2015-12-31_china_airports_part_3/)
sets by IATA.FAA code and will clean / fix missing data. We will create the
separate data set with the complement which will be fixed in next part.

In order to compare and fix missing or wrong data we are going to use different
tools :  

- `stringdist` from the package of the same name to check the distance between
two chains of characters
- the Google map API which give back elevation for a set of longitude / latitude


### Dependencies
We will need few standard packages as loaded bellow.

```{r libLoading, message=FALSE, warning=TRUE}
library(maptools)
library(reshape2)
library(grid)
library(dplyr)
library(ggplot2)
library(jsonlite)
library(Amelia)
library(OpenStreetMap)
library(stringdist)
library(grid)
```

### Loading part 3 objects

```{r loadPart1, echo=TRUE,eval=TRUE}
load(file = "./data/chinese-airports-part-3.Rda")
```


## Merging the two datasets by IATA.FAA code and extracting complements 
We will use IATA.FAA codes to identify the airports. First let's create subset of airport with and without IATA codes.

```{r isolateNoIATA}
OAwoIATA <- airportChinaOurairports[is.na(airportChinaOurairports$IATA.FAA),]
OAwithIATA <- airportChinaOurairports[!is.na(airportChinaOurairports$IATA.FAA),]
OFwoIATA <- airportChinaOpenflights[is.na(airportChinaOpenflights$IATA.FAA),]
OFwithIATA <- airportChinaOpenflights[!is.na(airportChinaOpenflights$IATA.FAA),]
```

Let's merge the two sub data sets with IATA code present. Then we isolate the remaining airport which have no match.
Regroup with airport without IATA and check if there are possible match by ICAO codes.

```{r mergingIATA}
## Merge by IATA, drop non matchin case
IATA <- merge(OAwithIATA, OFwithIATA, by = "IATA.FAA")
nrow(IATA)

## Extract OA airports not matching IATA code from OF
## And regroup with Airports without IATA
OAout <- OAwithIATA[!(OAwithIATA$IATA.FAA %in% IATA$IATA.FAA),]
OAout <- rbind(OAout, OAwoIATA)
## Extract OF airports not matching IATA code from OA
OFout <- OFwithIATA[!(OFwithIATA$IATA.FAA %in% IATA$IATA.FAA),]
OFout <- rbind(OFout, OFwoIATA)
## Check if there could be match by ICAO codes when present
sum(OAout[!is.na(OAout$ICAO),] %in% OFout[!is.na(OFout$ICAO),])
```

We have now three data sets to work on :  

- `IATA` which contains common airports by IATA code between openflight and ourairport data set. We need to define a strategy to select or merge their parameters if data is different (which elevation choose when different, which name, etc...)
- `OFout` airport from Openflight which have no match by IATA nor ICAO in ourairport data or no IATA/ICAO code.
- `OAout` airport from ourairport which have no match by IATA nor ICAO in openflights data or no IATA/ICAO code

For the two last data set, we merge them and we will have to find a way to collect missing data and add these to IATA data set in the later part of this report.

```{r actionPlan3}
# add missing categories to OFout
OFout$type.airport <- NA
OFout$scheduled.service <- NA
OFout$airport.wikipedia.link <- NA
OFout$region.name <- NA      
airports_2_check <- rbind(OFout, OAout)

missmap(IATA, main = "Missingness map for merged airport list by IATA.FAA Code")
missmap(airports_2_check, main = "Missingness map for remaining airports to be checked (not in both OA / OF sets")
```



## Checking and fixing IATA dataset
### Reorganising columns
We update column order to get a clearer view when comparing data.

```{r rearrangeIATA}
# reorder
IATA <- dplyr::select(IATA,
                      IATA.FAA, ICAO.x, ICAO.y,
                      name.x, name.y, city.x, city.y,
                      latitude.deg.x, latitude.deg.y,
                      longitude.deg.x, longitude.deg.y,
                      elevation.m.x, elevation.m.y,
                      category.x, category.y,
                      region.name, type.airport, scheduled.service, airport.wikipedia.link
) 
```

### Fixing ICAO codes  
Let's clean the ICAO codes, we will use the one from OurAiports. All missing 
values of OpenFlight are included in Ouraiports. Missing values of OurAirports
are not found in Openflight data set (see previous missingness map)


```{r full1}
IATA$ICAO <- IATA$ICAO.x
is.na(IATA[is.na(IATA$ICAO.x),]$ICAO.y)
IATA <- dplyr::select(IATA, -ICAO.x, -ICAO.y)
missmap(IATA)
```


### Fixing city name  
No missing data in city.y, we replace missing city.x by associated city.y.
There seems to be small typos in both data set (always for Chinese cities), 
we compare city names by calculating the distance between the character vectors.
We keep city.x which seems better.

```{r full4, message=TRUE}
IATA[is.na(IATA$city.x),]$city.x <- IATA[is.na(IATA$city.x),]$city.y 

dist <- stringdist(IATA$city.x, IATA$city.y)
hist(dist, breaks = 40)
dplyr::select(IATA[dist > 0,], city.x, city.y, name.x)

IATA$city <- IATA$city.x
IATA <- dplyr::select(IATA, -city.x, -city.y)
missmap(IATA)
```


### Fixing airport names  
No missing data in name.y or name.x, we need to compare both by string distance.
name.x is more detailed and better written. Except for Taoxian Airport.

```{r full5}
dist <- stringdist(IATA$name.x, IATA$name.y)
hist(dist, breaks = 40)
head(dplyr::select(IATA[dist > 0,], name.x, name.y))
IATA[IATA$name.x == "Taoxian Airport",]$name.x <- "Shenyang Taoxian International Airport"
IATA$name <- IATA$name.x
IATA <- dplyr::select(IATA, -name.x, -name.y)
missmap(IATA)
```


### Fixing longitude.deg, latitude.deg  
Values for longitude and latitude are different in the two data sets, we study 
in details the distance over 0.01 degree which represent roughly 500 meters. 

```{r mapAO2, warning=FALSE}
dist1 <- abs(IATA$longitude.deg.x - IATA$longitude.deg.y)
dist2 <- abs(IATA$latitude.deg.x - IATA$latitude.deg.y)

par(mfrow = c(1,2))
hist(log(dist1), breaks = 40)
hist(log(dist2), breaks = 40)
summary(dist1)
summary(dist2)


g <- ggplot() + theme_bw()
## airports with elevation
g <- g + geom_point(data = IATA[dist1 > 0.01 & dist2 > 0.01,], aes(x = longitude.deg.x, y = latitude.deg.x), col = "firebrick", pch = 3)
## airport with 0 m  as elevation in red
g <- g + geom_point(data = IATA[dist1 > 0.01 & dist2 > 0.01,], aes(x = longitude.deg.y, y = latitude.deg.y), col = "blue", pch = 1)
g <- g + geom_text(data = IATA[dist1 > 0.01 & dist2 > 0.01,], aes(x = longitude.deg.x, y = latitude.deg.x, label=IATA.FAA), size = 3, vjust = -1)
print(g)
```

The map above shows there are significant error between the two data sets, now
let's sample one case to check which data set is probably the most accurate.
We use for that the `OpenStreetMap` package to display side by side satellite 
images around the latitude and longitude coordinates of the airport in both data
set for the airport with largest distance.

```{r checkLocation, message=FALSE, fig.width=3, fig.height=3}
mapAirport <- function(
  long, lat,
  delta_lat=0.06,
  delta_lon=0.06,
  tiles=15, t="osm", #t="bing",
  title = "my title"
) {
  map = openmap(c(lat= lat-delta_lat,   lon= long-delta_lon),
                c(lat= lat+delta_lat,   lon= long+delta_lon),
                minNumTiles=tiles,type=t)
  plot(map)
}

toCheck <- dplyr::select(IATA[dist1 > 0.3 & dist2 > 0.3,],
                         IATA.FAA, ICAO, 
                         latitude.deg.x, longitude.deg.x,
                         latitude.deg.y,  longitude.deg.y)

for(i in 1){
  p1 <- mapAirport(
    long = toCheck[i,]$longitude.deg.x, lat = toCheck[i,]$latitude.deg.x,
    title = paste("OA", toCheck[i,]$IATA.FAA, "-", toCheck[i,]$ICAO)
  )
  print(p1)
  p2 <- mapAirport(
    long = toCheck[i,]$longitude.deg.y, lat = toCheck[i,]$latitude.deg.y,
    title = paste("OF:", toCheck[i,]$IATA.FAA, "-", toCheck[i,]$ICAO)
  )
  print(p2)
}
```

As a conclusion for this part, OA seems to be more accurate than OF regarding 
the airport longitude and latitude. We decide to keep OA data and drop OF data.
**note** we should confirm that openstreetmap satellite is accurate in China and 
there is not shift phenomenon. 

```{r fixLocation, message=FALSE}
IATA$latitude.deg <- IATA$latitude.deg.x
IATA$longitude.deg <- IATA$longitude.deg.x
IATA <- dplyr::select(IATA, -latitude.deg.x, -latitude.deg.y, -longitude.deg.x, -longitude.deg.y)
```

Nevertheless as shown in the case of ZLXH airport the data of OA seems 
inaccurate. Further check is probably needed. **note 2** Comparison is done here
manually, we will see in another post how to make this detection using machine
learning algorithms.


### Fixing categories
They are same in both data sets except the following. We decide to keep OF
categories.

```{r fixCat}
## mismatch in category
IATA[!(IATA$category.x == IATA$category.y),]$name
IATA$category <- IATA$category.y
IATA <- dplyr::select(IATA, -category.x, -category.y)
missmap(IATA)
```




### Fixing elevation.m data  
As we have decided to drop OF latitude and longitude we should also drop the
elevation from OF data set.

```{r full3}
IATA$elevation.m <- IATA$elevation.m.x
IATA <- dplyr::select(IATA, -elevation.m.y, -elevation.m.x)
```

We have two actions to check and fix elevation data : check if the elevation is
consistent with latitude and longitude and find missing elevation data.

We can use for this purpose the jawq API which returns in JSON format the
elevation value for a given set of latitude / longitude.

```{r checkElevation, cache=TRUE}
# function which return evelevation based on longitude / latitude
# https://www.jawg.io/docs/apidocs/elevation/#request-query-parameters
api_key <- "lubGnrGVymuYM1DU9Dqu8grqy9sMmvJhd25RPHmandKU0DfTuGcjbz3KgDs4vPIk"

getElevation <- function(lat,long) {
  lat <- as.numeric(lat)
  long <- as.numeric(long)
  # elevation API
  baseUrl <- "https://api.jawg.io/elevations?locations="
  Url <- paste(baseUrl, lat, "," , long, "&access-token=", api_key, sep = "")
  jsonData <- fromJSON(Url)
  Sys.sleep(0.1)
  elevation <- jsonData$elevation
  elevation
  #Url
}
## get elevation for all aiports
#IATA$elevationCheck <- 1
elevationCheck <- apply(IATA, 1, function(x) getElevation(x[9],x[10]))
```


```{r}
IATA$elevationCheck <- elevationCheck
```



We are now going to check the gap between Google data and OA data set.

```{r replaceElevation}
## replace missing elevation by google values
IATA[is.na(IATA$elevation.m),]$elevation.m <- IATA[is.na(IATA$elevation.m),]$elevationCheck
dist <- (abs(IATA$elevation.m - IATA$elevationCheck))
hist(dist, breaks = 40)
summary(dist)
dplyr::select(IATA[dist > 30,], IATA.FAA, ICAO, elevation.m, elevationCheck, longitude.deg, latitude.deg)
```

We have `r sum(dist > 30)` airports with more than 30 meters of elevation error.
After checking, we find out that Jawg data is more accurate and use it for 
these airports.

```{r}

IATA[dist > 30,]$elevation.m <- IATA[dist > 30,]$elevationCheck
IATA <- dplyr::select(IATA, -elevationCheck)
missmap(IATA)
```



We use information from [http://www.geoplaner.com/#Coordinate-Converter] for
setting


## Saving objects for next part
Let's now save the different objects to be reused in the next part of this post.

```{r saveObjects, cache=FALSE}
save( list = c("IATA", 
               "airports_2_check"
),
file = "./data/chinese-airports-part-4.Rda")
```




## Code information

### Source code

The source code of this post is available on [github]()

### Session information

```{r}
sessionInfo()
```




