---
title: "The new french election geography"
date: "2019-09-21T13:39:46+02:00"
author: "Datapleth.io"
tags: ["choropleth", "france"]
categories: ["geography","france"]
banner: "blog/2019-09-21_the_french_new_election_geography_files/figure-html/electionChoropleth-1.png"
output: html_document
---

The last European elections were on June 26th, 2019. One of the major
observation was the increase of extreme right votes, and more particularily
the shift toward extreme right of areas which used to be left voters. In this 
post we are going to visualize the new shape of France.



```{r loadLibraries, message=FALSE, warning=FALSE}
library(data.table)
library(knitr)
library(kableExtra)
library(dplyr)
library(geojson)
library(broom)
library(ggplot2)
library(ggthemes)
library(rmapshaper)
```

## Election results

We need first to download the official results from
[data.gouv.fr](https://www.data.gouv.fr/en/datasets/resultats-des-elections-europeennes-2019/)

```{r getData, message=FALSE, warning=FALSE}
results_uri <- "https://www.data.gouv.fr/en/datasets/r/35170deb-e5f3-4e79-889f-5b9a3f547742"
election_results <- data.table::fread(
  results_uri, encoding = "Latin-1", dec= ","
)
```

In the election result file we don't have the insee code of the "communes" (an
equivalent of counties in USA), we have to build that based on department number
and county code. This is working well except for data outside France mainland.

```{r getInsee, message=FALSE, warning=FALSE}
## ZZ for other countries -> 99
election_results[ , insee := as.numeric(`Code du département`)]
election_results[ 
  `Code du département` == "ZZ",
  insee := 99
  ]
election_results[ 
  `Code du département` %in% c("ZX","ZS","ZM","ZD","ZC","ZB","ZA"),
  insee := 97
  ]
election_results[ 
  `Code du département` %in% c("ZW","ZP","ZN"),
  insee := 98
  ]
election_results[, insee := as.character(insee*1000 + `Code de la commune`)]
election_results[ nchar(insee) == 4, insee := paste0("0",insee)]
```


The format of the data is really awfull, headers are not covering all columns.
We first select the interesting columns and rename. We'll use first the results
of the main lists. Communes data are store in columns 1 to 18, then each
list results are stored in 7 columns.

```{r extractList, message=FALSE, warning=FALSE}
extract_list <- function(data,list_id = 23){
  # compute list index and extract it
  idx <-  c(1:18, (19+(list_id -1)*7):(19+(list_id -1)*7+6),257)
  sub_data <- data[ , .SD, .SDcols=idx]
  names_clean <- c(
    "departement_id", #[1] "Code du département"   
    "deparement_name", #[2] "Libellé du département"
    "commune_id", #[3] "Code de la commune"    
    "commune_name", #[4] "Libellé de la commune" 
    "inscrits", #[5] "Inscrits"              
    "abstentions", #[6] "Abstentions"           
    "absentions_perc", #[7] "% Abs/Ins"             
    "votants", #[8] "Votants"               
    "votants_perc", #[9] "% Vot/Ins"             
    "blancs", #[10] "Blancs"                
    "blancs_perc", #[11] "% Blancs/Ins"          
    "blancs_perc_votants", #[12] "% Blancs/Vot"          
    "nuls", #[13] "Nuls"                  
    "nuls_perc", #[14] "% Nuls/Ins"            
    "nuls_perc_votants", #[15] "% Nuls/Vot"            
    "exprimes", #[16] "Exprimés"              
    "exprimes_perc", #[17] "% Exp/Ins"             
    "exprimes_perc_votants", #[18] "% Exp/Vot"             
    "liste_id", #[19] "N°Liste"               
    "liste_shortname", #[20] "Libellé Abrégé Liste"  
    "liste_name", #[21] "Libellé Etendu Liste"  
    "tete_de_liste", #[22] "Nom Tête de Liste"     
    "nb_voix", #[23] "Voix"                  
    "nb_voix_perc", #[24] "% Voix/Ins"            
    "nb_voix_perc_exprimes", #[25] "% Voix/Exp" 
    "code_insee"
  )
  setnames(sub_data, names_clean)
  return(sub_data)
}

election_results_rn <- extract_list(data = election_results, list_id = 23)
election_results_lrem <- extract_list(data = election_results, list_id = 5)
election_results_lfi <- extract_list(data = election_results, list_id = 1)
election_results_ps <- extract_list(data = election_results, list_id = 12)
election_results_lr <- extract_list(data = election_results, list_id = 29)
election_results_eelv <- extract_list(data = election_results, list_id = 30)

election_results_sub <- rbind(
  election_results_rn,
  election_results_lrem,
  election_results_lfi,
  election_results_ps,
  election_results_lr,
  election_results_eelv
  )
```


We have now a clean table with data for the main parties lists.

```{r}
knitr::kable(head(election_results_sub)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "100%")
```


## France 'communes' shapefiles.

We are getting the shapefile stored as geojson formats directly on 
[data.gouv.fr](https://www.data.gouv.fr/en/datasets/decoupage-administratif-communal-francais-issu-d-openstreetmap/#_), however they are stored in zip and we store a copy
in our cloud storage.



```{r getFranceSp, message=FALSE, warning=FALSE}
## url of the geojson file downloaded and unzipped from data.gouv.fr
france_sp_uri <- "https://data.datapleth.io/ext/france/spatial/communes-simple/communes-20190101.json"

## let's read directly the file
france_communes <- geojsonio::geojson_read(france_sp_uri,  what = "sp")
```

Let's filter the results of France mainland only

```{r filterShape, message=FALSE, warning=FALSE}
communes_lim <- france_communes[ ! substr(france_communes@data$insee,1,2)  %in% c(
    "97","98","99"
    ), ]
```

This is a quite large and detailled file, we are simplifying...

```{r simplifyMap, message=FALSE, warning=FALSE}
communes_lim <- rmapshaper::ms_simplify(communes_lim)
```



Let's have a look on this map with the communes areas.

```{r message=FALSE, warning=FALSE}
# Fortify the data AND keep trace of the commune code.
communes_lim_fortified <- broom::tidy(communes_lim, region = "insee")
```

```{r plotBaseMap, message=FALSE, warning=FALSE}
# Now I can plot this shape easily as described before:
ggplot() +
  geom_polygon(data = communes_lim_fortified, 
               aes( x = long, y = lat, group = group),
               fill="white", 
               color="grey", size = 0.2
               ) +
  theme_void() +
  coord_map()
```



## Election choropleth

We have election results per communes and spatial file per communes. Let's merge
both datasets to plot a choropleth.

```{r mergeDataShape, message=FALSE, warning=FALSE}
communes_results <- merge(
  x = election_results_sub
  , y = communes_lim_fortified
  , by.x = "code_insee"
  , by.y = "id"
  , allow.cartesian=TRUE
)
```

Finaly we can plot the results with one map per list.



```{r electionChoropleth, message=FALSE, warning=FALSE}
p <- ggplot() +
  geom_polygon(
    data = communes_results, 
    aes(fill = nb_voix_perc_exprimes, x = long, y = lat, group = group),
    color="grey", size = 0.02
  ) +
  scale_fill_viridis_c(option = "A", direction = -1) +
  facet_wrap(facets = . ~ liste_shortname) +
  coord_map() +
  theme_tufte() +
  theme(
    axis.line=element_blank()
    , axis.text=element_blank()
    , axis.ticks=element_blank()
    , axis.title=element_blank()
    ) +
  ggtitle("European election - 2019 - Limousin (France)") +
  labs(fill = "% of votes")
p
```


```{r rayshading, eval=FALSE, warning=FALSE, include=FALSE}
#library(rayshader)
#rayshader::plot_gg(p,multicore=TRUE,width=5,height=5,scale=250)
#rgl::view3d(theta = 45, phi = 30, zoom = 0.4, fov = 50)
```



## References

https://www.r-graph-gallery.com/327-chloropleth-map-from-geojson-with-ggplot2.html