---
title: "The french yield curve problem? (part 1)"
date: "2019-08-28T13:39:46+02:00"
author: "Datapleth.io"
tags: ["3d", "plotly"]
categories: ["economy"]
banner: "img/illustrations/datapleth-yield-curve.png"
output: html_document
---


Following our [previous]() post on the USA yield curve inversion, we are going 
to evaluate the situation in other countries, starting with France.

## Preparation

Let's as usual load the libraries we need.

```{r message=FALSE, warning=FALSE}
library(xml2)
library(dplyr)
library(data.table)
library(plotly)
library(knitr)
library(kableExtra)
Sys.setlocale("LC_ALL","C")
```


## Getting and cleaning the data for France

Data source for France data is French central bank, known as "Banque de France" (see [here](https://www.banque-france.fr/statistiques/taux-et-cours/taux-indicatifs-des-bons-du-tresor-et-oat). They provide the data as HTML and CSV as one URL with all 
historical values.

```{r getFranceData1, message=FALSE, warning=FALSE}
url_france <- "http://webstat.banque-france.fr/fr/downloadFile.do?id=5385691&exportType=csv"
## Get column information, including name
france_features <- read.csv(
    file = url_france,
    header = FALSE,
    nrows = 6,
    stringsAsFactors = FALSE,
    sep = ";")
## Get Data
all_france_data_raw <- read.csv(
    file = url_france,
    skip = 6,
    header = FALSE,
    stringsAsFactors = FALSE,
    sep = ";")
```

Once this is done, we convert the data.frame as data.table and we check an
extract. 

```{r getFranceData2, message=FALSE, warning=FALSE}
data.table::setDT(all_france_data_raw) 
data.table::setnames(all_france_data_raw, make.names(as.character(france_features[2,])))
knitr::kable(head(all_france_data_raw)) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

We change now the features names to get a similar structure as for USA data.
We will also have to clean the table as numeric separator are commas and 
missing values are dashes.

```{r cleanFranceData}
## a function to convert "-0,371" as "-0.371" and "-" as NA
convert2string <- function(x){
    as.numeric(
        gsub(
            pattern = ",",
            replacement = ".",
            x = x
        )
    )
}

## We use data table "way" to convert all numerical features
all_france_data <- all_france_data_raw[ ,
                                        lapply(.SD, convert2string) ,
                                        .SDcols = names(all_france_data_raw) %like% "FM"
                                        ]
knitr::kable(head(all_france_data)) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


```{r}
## add the date back
all_france_data <- cbind(
    all_france_data_raw[ , Code.s..rie..],
    all_france_data)

## Compute dates in correct locale
all_france_data[ , NEW_DATE := as.Date(
    as.POSIXlt(x = all_france_data_raw$Code.s..rie..,
               format = "%d/%m/%Y")
)
]
all_france_data[ , V1 := NULL ]
knitr::kable(head(all_france_data)) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

```{r}
## We change the names
old_names <- c(
    "NEW_DATE",
    "FM.D.FR.EUR.FR2.BB.FR1MT_RR.YLD",
    "FM.D.FR.EUR.FR2.BB.FR3MT_RR.YLD",
    "FM.D.FR.EUR.FR2.BB.FR6MT_RR.YLD",
    "FM.D.FR.EUR.FR2.BB.FR9MT_RR.YLD",
    "FM.D.FR.EUR.FR2.BB.FR1YT_RR.YLD",
    "FM.D.FR.EUR.FR2.BB.FR2YT_RR.YLD",
    "FM.D.FR.EUR.FR2.BB.FR5YT_RR.YLD",
    "FM.D.FR.EUR.FR2.BB.FR10YT_RR.YLD",
    "FM.D.FR.EUR.FR2.BB.FR30YT_RR.YLD"
)
new_names <- c(
    "NEW_DATE",
    "BC_1MONTH",
    "BC_3MONTH",
    "BC_6MONTH",
    "BC_9MONTH",
    "BC_1YEAR",
    "BC_2YEAR",
    "BC_5YEAR",
    "BC_10YEAR",
    "BC_30YEAR"
)
data.table::setnames(
    x = all_france_data, 
    old = old_names, 
    new = new_names)
```


Finally we create a data set in long format.

```{r message=FALSE, warning=FALSE}
all_france_data_long <- data.table::melt(
    all_france_data,
    measure.vars = c(
        "BC_1MONTH", "BC_3MONTH", "BC_6MONTH", "BC_9MONTH",
        "BC_1YEAR", "BC_2YEAR", "BC_5YEAR", 
        "BC_10YEAR", "BC_30YEAR"),
    variable.name = "horizon",
    value.name = "rate")
```

And we convert horizon to numerical values in months, add a data column as 
numeric.

```{r message=FALSE, warning=FALSE}
all_france_data_long[ horizon == "BC_1MONTH", horizon := "1"]
all_france_data_long[ horizon == "BC_3MONTH", horizon := "3"]
all_france_data_long[ horizon == "BC_6MONTH", horizon := "6"]
all_france_data_long[ horizon == "BC_9MONTH", horizon := "9"]
all_france_data_long[ horizon == "BC_1YEAR", horizon := "12"]
all_france_data_long[ horizon == "BC_2YEAR", horizon := "24"]
all_france_data_long[ horizon == "BC_5YEAR", horizon := "60"]
all_france_data_long[ horizon == "BC_10YEAR", horizon := "120"]
all_france_data_long[ horizon == "BC_30YEAR", horizon := "360"]
all_france_data_long[, horizon := as.numeric(as.character(horizon))]
save(all_france_data_long, file = "./data/all_france_rate.Rda")
```

We make a quick plot, it's always usefull to explore the data. 

```{r message=FALSE, warning=FALSE}
g <- ggplot(all_france_data_long)
g <- g + geom_point(aes(x = NEW_DATE,
                        y = rate,
                        col = horizon),
                    alpha = 0.5, size = 0.5)
g <- g + facet_grid(facets =  horizon ~ . )
g
```


## Plotting the 3d yield curve

There are several alternatives to plot 3d surfaces in R but to make it
interactive, we choose the `plotly` package. We have to move back the data in
a long format.

For the surface color, we will plot it as the ration to rate with 3month horizon
as reference. Thus we create a new matrix with the calculation and deal with 
special values when 3 months rate is 0.


```{r message=FALSE, warning=FALSE}
## reshape de data to get a matrix (for plotly)
d <- copy(all_france_data_long)
d <- data.table::dcast(d, NEW_DATE ~ horizon, value.var = "rate")
d[ , NEW_DATE := NULL]

## compute ratio to 3 month rate
c <- copy(d)
setnames(c, make.names(names(c)))
c[ , ':=' (
    X1 = (X1-X3) / abs(X3),
    X6 = (X6-X3) / abs(X3),
    X9 = (X9 - X3) / abs(X3),
    X12 = (X12 - X3) / abs(X3),
    X24 = (X24 - X3) / abs(X3),
    X60 = (X60 - X3) / abs(X3),
    X120 = (X120 - X3) / abs(X3),
    X360 = (X360 - X3) / abs(X3)
)]
## deal with zero values
c[ X3 == 0 | is.na(X3), ':=' (
    X1 = NA,
    X6 = NA,
    X9 = NA,
    X12 = NA,
    X24 = NA,
    X60 = NA,
    X120 = NA,
    X360 = NA
)]
c[ , X3 := 1 ]
```

We are now ready to plot. We choose the color theme of Dark2 palette (green & 
orange).


```{r plot3d, message=FALSE, warning=FALSE, out.width='100%'}
p <- plot_ly(
    x = sort(unique(all_france_data_long$NEW_DATE)),
    y = sort(unique(all_france_data_long$horizon)),
    z = t(as.matrix(d)),
    type = "surface",
    surfacecolor = t(as.matrix(c)),
    cmin = -1,
    cmax = +1,
    colorscale = list(
        list(
            0,
            "rgb(215, 95, 2)"
        ),
        list(
            0.5,
            "rgb(231, 245, 255)"
        ),
        list (
            1,
            "rgb(25, 155, 115)"
        )
    ),
    colorbar = list(
        title='ratio to<br>3 month<br>yield',
        side = 'bottom',
        thickness='10',
        xpad = 5,
        y = 0.8
    ),
    lighting = list(
        ambient = 0.8,
        diffuse = 0.8,
        specular = 0.2,
        roughness = 0.8,
        fresnel = 0.2
    ),
    opacity = 0.9,
    hoverlabel = list(
        bgcolor = "rgb(255, 255, 255)"
    )
) %>%
    plotly::layout(
        #title = "3D yield curve",
        width = 800,
        height = 500,
        scene=list(
            xaxis=list(title="date"),
            yaxis=list(title="horizon"),
            zaxis=list(title="rate"),
            aspectmode = "manual",
            aspectratio = list(x=4,y=2,z=1.3),
            camera = list(
                eye = list(x = 3, y = -3, z = 0.3 ),
                center = list( x = 0.8, y = 0, z = 0)
            )
        )
    ) %>% 
    config(displayModeBar = F) 
p
```

## References

From The New York Times :

- https://www.nytimes.com/2019/08/15/upshot/inverted-yield-curve-bonds-football-analogy.html
- https://www.nytimes.com/interactive/2015/03/19/upshot/3d-yield-curve-economic-growth.html?module=inline

Other R works :

- https://www.r-bloggers.com/plotting-the-evolution-of-the-u-s-treasury-yield-curve/
- https://plot.ly/~PlotlyTutorials/326/us-treasury-yield-curve/#/
- https://gist.github.com/timelyportfolio/4da9d6b6c89cce26effabccca30124dd
- https://stefan.angrick.me/a-3d-view-of-the-us-treasury-yield-curve


