---
title: "The unicorn bubble - Donuts or Waffles ?"
date: "2019-12-08T13:39:46+02:00"
author: "Datapleth.io"
tags: ["NYSE", "WSJ", "waffle"]
categories: ["finance", "dataviz"]
banner: "img/illustrations/waffle-ipo.png"
output: html_document
---



## How Silicon Valley feels right now ?

Few days ago the Wall Stree Journal (@WSGGraphics) published a tweet about 
Unicorns loosing value between their last private-market valuation and their
IPO price. 

In this post, we will replicate their visuals and we will demonstrate the 
problem of using donut like charts. We will propose an alternative using
waffle charts.

They illustrated the analysis with some circular charts. The problem with such
visuals is that human eye is very bad ad comparing surfaces when they are 
represented by disks.

WSJ objective was to present a comparison of these companies valuation, 
In addition, WSJ wanted to illustrate the loss of value during IPO
using two colors. If we look at the comparison between Wework
and Juul Labs, this is striking. The two half disks looks very similar in term
of diameters, but there is a `r round(47/38*100-100)`% gap ! WeWork valuation is
47 Billions USD vs. Juul Labs' is 38 billions USD,

In addition, the loss of value of Juul Labs looks relatively small as the gap 
between the inner disk and the outside radius is close but in reality, this
company lost `r round(100-24/38*100)`% !


<center>
<blockquote class="twitter-tweet"><p lang="en" dir="ltr">‘We are all adjusting our eyes and no one has any idea how the rest of the night is going to go. That’s how Silicon Valley feels right now.’<a href="https://t.co/spPPAD7ggZ">https://t.co/spPPAD7ggZ</a> via <a href="https://twitter.com/WSJ?ref_src=twsrc%5Etfw">@WSJ</a> <a href="https://t.co/C33VDREVUH">pic.twitter.com/C33VDREVUH</a></p>&mdash; WSJ Graphics (@WSJGraphics) <a href="https://twitter.com/WSJGraphics/status/1199347440559648774?ref_src=twsrc%5Etfw">November 26, 2019</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
</center>


## Let's replicate WSJ visuals

As a first step, we will reproduce the WSJ plots using the R package `ggforce`.
This package is adding interesting functionnalites to `ggplot2`, in particular
the possibility to draw circles. The half disc surface represent the company
valuation, the larger as before IPO and the smaller in grey, after IPO.

The formula to compute the radius of each circle can be deduced from the surface
of the half disk.

$$ S = \frac{\pi . r² }{2} $$  

Which makes :

$$r = \sqrt{\frac{2.S}{\pi}} $$



```{r makeWJSplot, message=FALSE, warning=FALSE}
library(ggplot2)
library(ggforce)

# color 
wjs <- c("#ad3976", "#eeeff1")

# defintion of data
unicorns <- data.frame(
    start = rep(-pi/2,4)
    , end = rep(pi/2,4)
    , valuation = c(47, 8, 38, 24)
    , type = c("IPO","Private","IPO","Private")
    , unicorn = factor(
        c("WeWork","WeWork", "Juul Labs", "Juul Labs")
        , levels = c("WeWork", "Juul Labs")
    )
)

loss <- data.frame(
    x = c(0,0)
    , y = c(-0.3,-0.3)
    , loss = c("-$39 B","-$14 B")
    , unicorn = c("WeWork", "Juul Labs")
)

# Behold the arcs
ggplot(unicorns) +
    geom_arc_bar(
        aes(
            x0 = 0, y0 = 0, r0 = 0, r = sqrt(valuation/pi*2),
            start = start, end = end, fill = type), col = "#eeeff1"
    ) + 
    coord_fixed() +
    geom_text(data = loss, aes(x = x,y = y,label = loss), col = wjs[1], fontface = "bold") +
    theme_void() +
    scale_fill_manual(values = wjs) +
    scale_color_manual(values = wjs) +
    facet_wrap(facets = . ~ unicorn) + theme( legend.position = "none")

```


It's a nice visual, we can see clearly the hug frop of value of the failed 
WeWork IPO. However it's very difficult to compare the pre-IPO value of Juul 
Labs with the one of WeWork if we don't add the legend (as WSJ did it). We
proposes bellow an alternative based on waffle charts where one square 
represents 1 billion USD. It makes the comparison between the two companies 
easier, as well as the drop of value from Juul Labs more visible.


```{r waffle, message=FALSE, warning=FALSE}
library(waffle)
ww <- c(`IPO Value`=(8), `Loss from
private valuation`=39)
jl <- c(`IPO Value`=(24), `Loss from
private valuation`=14)
iron(
    waffle(ww, rows=3, size=1, colors=c(wjs[2], wjs[1]), 
           title="WeWork", 
           xlab="1 b$ per square",
           pad = 1),
    waffle(jl, rows=3, size=1, colors=c(wjs[2], wjs[1]), 
           title="Juul Labs", 
           xlab="1 b$ per square",
           pad = 4)
)
```


## Conclusions

Are you Donuts of Waffles ?


## References

(https://www.wsj.com/articles/silicon-valley-adjusts-to-new-reality-as-100-billion-evaporates-11574764205?shareToken=st69f61faedfb349d3b584f9bc1dcf58db&mod=e2twg)