---
title: "The french new election geography"
date: "2019-09-21T13:39:46+02:00"
author: "Datapleth.io"
tags: ["3d", "plotly"]
categories: ["economy"]
banner: "img/illustrations/datapleth-yield-curve-france.png"
output: html_document
---

The last European elections were on June 26th, 2019. One of the major
observation was the increase of extreme right votes, and more particularily
the shift toward extreme right of areas which used to be left voters. In this 
post we are going to visualize the new shape of France.



```{r loadLibraries, message=FALSE, warning=FALSE}
library(data.table)
library(knitr)
library(kableExtra)
library(dplyr)
library(geojson)
library(broom)
library(ggplot2)
library(ggthemes)
```

## Election results

We need first to download the official results from
[data.gouv.fr](https://www.data.gouv.fr/en/datasets/resultats-des-elections-europeennes-2019/)

```{r getData}
results_uri <- "https://www.data.gouv.fr/en/datasets/r/35170deb-e5f3-4e79-889f-5b9a3f547742"
election_results <- data.table::fread(
  results_uri, encoding = "Latin-1", dec= ","
)
```

In the election result file we don't have the insee code of the county, we have
to build that based on department number and county code. This is working well
except for data outside France mainland.

```{r getInsee}
## ZZ for other countries -> 99
election_results[ , insee := as.numeric(`Code du département`)]
election_results[ 
  `Code du département` == "ZZ",
  insee := 99
  ]
election_results[ 
  `Code du département` %in% c("ZX","ZS","ZM","ZD","ZC","ZB","ZA"),
  insee := 97
  ]
election_results[ 
  `Code du département` %in% c("ZW","ZP","ZN"),
  insee := 98
  ]
election_results[, insee := as.character(insee*1000 + `Code de la commune`)]
election_results[ nchar(insee) == 4, insee := paste0("0",insee)]
```



The format of the data is really awfull, headers are not covering all columns.
We first select the interesting columns and rename. We'll use first the results
of the main extreme right party (Rassemblement National) which are in column
`V61` to `V68`. County (Commune) data are store in columns 1 to 18, then each
list results are stored in 8 columns.

```{r extractList}
extract_list <- function(data,list_id = 23){
  # compute list index and extract it
  idx <-  c(1:18, (19+(list_id -1)*7):(19+(list_id -1)*7+6),257)
  sub_data <- data[ , .SD, .SDcols=idx]
  names_clean <- c(
    "departement_id", #[1] "Code du département"   
    "deparement_name", #[2] "Libellé du département"
    "commune_id", #[3] "Code de la commune"    
    "commune_name", #[4] "Libellé de la commune" 
    "inscrits", #[5] "Inscrits"              
    "abstentions", #[6] "Abstentions"           
    "absentions_perc", #[7] "% Abs/Ins"             
    "votants", #[8] "Votants"               
    "votants_perc", #[9] "% Vot/Ins"             
    "blancs", #[10] "Blancs"                
    "blancs_perc", #[11] "% Blancs/Ins"          
    "blancs_perc_votants", #[12] "% Blancs/Vot"          
    "nuls", #[13] "Nuls"                  
    "nuls_perc", #[14] "% Nuls/Ins"            
    "nuls_perc_votants", #[15] "% Nuls/Vot"            
    "exprimes", #[16] "Exprimés"              
    "exprimes_perc", #[17] "% Exp/Ins"             
    "exprimes_perc_votants", #[18] "% Exp/Vot"             
    "liste_id", #[19] "N°Liste"               
    "liste_shortname", #[20] "Libellé Abrégé Liste"  
    "liste_name", #[21] "Libellé Etendu Liste"  
    "tete_de_liste", #[22] "Nom Tête de Liste"     
    "nb_voix", #[23] "Voix"                  
    "nb_voix_perc", #[24] "% Voix/Ins"            
    "nb_voix_perc_exprimes", #[25] "% Voix/Exp" 
    "code_insee"
  )
  setnames(sub_data, names_clean)
  return(sub_data)
}

election_results_rn <- extract_list(data = election_results, list_id = 23)
election_results_lrem <- extract_list(data = election_results, list_id = 5)
election_results_lfi <- extract_list(data = election_results, list_id = 1)
election_results_ps <- extract_list(data = election_results, list_id = 12)
election_results_lr <- extract_list(data = election_results, list_id = 29)
election_results_eelv <- extract_list(data = election_results, list_id = 30)

election_results_sub <- rbind(
  election_results_rn,
  election_results_lrem,
  election_results_lfi,
  election_results_ps,
  election_results_lr,
  election_results_eelv
  )
```


We have now a clean table with data for Rassemblement National list and LREM.

```{r}
knitr::kable(head(election_results_sub)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "100%")
```


## France 'communes' (counties?)

We are getting the shapefile stored as geojson formats directly on 
[data.gouv.fr](https://www.data.gouv.fr/en/datasets/decoupage-administratif-communal-francais-issu-d-openstreetmap/#_).



```{r getFranceSp, message=FALSE, warning=FALSE}
## url of the geojson file downloaded and unzipped from data.gouv.fr
france_sp_uri <- "https://data.datapleth.io/ext/france/spatial/communes-simple/communes-20190101.json"

## let's read directly the file
france_communes <- geojsonio::geojson_read(france_sp_uri,  what = "sp")
```

```{r filterShape}
## let's filter only the region of Limousin
communes_lim <- france_communes[ 
  substr(france_communes@data$insee,1,2)  %in% c(
    "87","19","23"
    ), 
  ]
```



```{r message=FALSE, warning=FALSE}
# I need to fortify the data AND keep trace of the commune code! (Takes ~2 minutes)
communes_lim_fortified <- broom::tidy(communes_lim, region = "insee")

# Now I can plot this shape easily as described before:
ggplot() +
  geom_polygon(data = communes_lim_fortified, aes( x = long, y = lat, group = group), fill="white", color="grey") +
  theme_void() +
  coord_map()
```



## Election choropleth

We have election results per communes and spatial file per communes. Let's merge
both datasets to plot a choropleth.

```{r}
communes_results <- merge(
  x = election_results_sub
  , y = communes_lim_fortified
  , by.x = "code_insee"
  , by.y = "id"
  , allow.cartesian=TRUE
)
```





```{r}
p <- ggplot() +
  geom_polygon(
    data = communes_results, 
    aes(fill = nb_voix_perc_exprimes, x = long, y = lat, group = group),
    color="grey", size = 0.2
  ) +
  scale_fill_viridis_c(option = "A", direction = -1) +
  facet_wrap(facets = . ~ liste_shortname) +
  coord_map() +
  theme_tufte() +
  theme(
    axis.line=element_blank()
    , axis.text=element_blank()
    , axis.ticks=element_blank()
    , axis.title=element_blank()
    ) +
  ggtitle("European election - 2019 - Limousin (France)") +
  labs(fill = "% of votes")
p
```


```{r}
#library(rayshader)
#rayshader::plot_gg(p,multicore=TRUE,width=5,height=5,scale=250)
#rgl::view3d(theta = 45, phi = 30, zoom = 0.4, fov = 50)
```



## References

https://www.r-graph-gallery.com/327-chloropleth-map-from-geojson-with-ggplot2.html